Flowers text analysis

### 2022-12-07 code/annotation_analysis.Rmd
wrangling adding zeros and code cleanup

```{r}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName, condition)

annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
  filter(!is.na(span)) |> 
  mutate(numchar=str_length(span),
         numword=str_count(span, "\\W+") + 1) |> 
  left_join(d.games) |> 
  mutate(flower=str_c(referent,color),
         player=str_c(gameId, name)) |> 
  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)
```

@@

message("total utterances")
annotated |> tally()

@@

annotated |> group_by(gameId,condition) |> tally() |> pull(n) |> summary()

message("Utterances / flower")
annotated |> group_by(flower) |> tally() |> pull(n) |> summary()

@@

sum_words <- annotated %>% group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
                                           sum_word=sum(numword))

@@

        words <- ggplot(annotated,aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.1)+ geom_smooth()+
          labs(y="Words per reference expression", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")

        utts <- annotated |> group_by(gameId, trialNum, condition) |> tally() |> 
  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.02)+geom_smooth()+
          labs(y="Reference expressions  per trial per game", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")

@@

@@

annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> 
  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.2)+geom_smooth()+
          labs(y="Distinct flowers referenced per trial", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")
```

```{r}
annotated |> group_by(gameId, trialNum, condition, player) |> select(flower) |> unique() |> tally() |> 
  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.1)+geom_smooth()+
          labs(y="Distinct flowers referenced per player per trial", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")
```

@@

```{r, include=F,}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName, condition)

@@

 sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv"))|> 
    left_join(d.players) |> left_join(d.games) |> mutate(seen="game") |> 
   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()

sbert_post_test <- read_csv(here(annotation_path,"post_test.csv")) |> left_join(d.games) |> 
  mutate(trialNum=25) |> 
  select(playerId, gameId, condition, text,referent=num, color, trialNum, seen)%>% bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData')))  %>% as_tibble()

sbert_all <- sbert_single |> union(sbert_post_test)
F_mat <- sbert_all %>% select(starts_with("...")) %>% as.matrix() #Features
M_mat <- sbert_all %>% 
  select(-starts_with("...")) %>% 
  mutate(feature_ind=row_number(),
         sixth=case_when(
           trialNum < 4 ~ 1,
          trialNum < 8 ~ 2,
          trialNum < 12 ~ 3,
          trialNum < 16 ~ 4,
          trialNum < 20 ~ 5,
          trialNum < 24 ~ 6,
          trialNum > 24 ~ 7
  ))
```

>>>>

```{r}
d.games <- read_csv(here(read_data_path, "games.csv")) |> 
                        mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  filter(chatEnabled) |> 
  select(gameId,conditionName, condition)

n_players = 3
n_rounds = 24

d.rounds <- read_csv(here(read_data_path,"rounds.csv"))
d.players <- read_csv(here(read_data_path, "players.csv")) |> select(playerId,name)
game_completion <- d.rounds %>% 
  group_by(gameId) %>% 
  tally() %>% 
  mutate(game_complete = n==n_players*n_rounds) %>%
  select(gameId, game_complete)

d.games.final <- d.games %>% 
  left_join(game_completion) %>% 
  replace_na(list(game_complete = FALSE, n_trials = 0)) |> 
  filter(game_complete)

d.rounds.final <-  d.rounds %>% 
  inner_join(d.games.final) %>% 
  filter(conditionName != "coopCartel",
         game_complete) |> 
  left_join(d.players) |> 
  #left_join(d.games) |> 
  select(gameId,trialNum, name, condition)

annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
  filter(!is.na(normalized_span)) |> 
  inner_join(d.games.final) |> 
  mutate(numchar=str_length(span),
         numword=ifelse(numchar==0,0,str_count(span, "\\W+") + 1)) |> 
  mutate(flower=str_c(referent,color),
         player=str_c(gameId, name)) |> 
  select(gameId, player,name, flower, trialNum, pResp, normalized_span,color,numchar, numword, condition)

@@

message("total utterances")
annotated |> filter(numword>0) |>  tally()

@@

annotated |> filter(numword>0) |> group_by(gameId,condition) |> tally() |> pull(n) |> summary()

message("Utterances / flower")
annotated |> filter(numword>0) |> group_by(flower) |> tally() |> pull(n) |> summary()

@@

sum_words <- annotated %>% filter(numword>0) |>  group_by(trialNum, gameId,condition) %>% summarize(sum_char=sum(numchar),
                                           sum_word=sum(numword))

@@

        words <- ggplot(annotated |> filter(numword>0),aes(x=trialNum,y=numword, color=condition))+geom_jitter(alpha=.1)+ geom_smooth()+
          labs(y="Words per reference expression", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")

        utts <- annotated |> filter(numword>0) |> group_by(gameId, trialNum, condition) |> tally() |> 
  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(alpha=.02)+geom_smooth()+
          labs(y="Reference expressions  per trial per game", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")

@@

```{r}
possible <- d.rounds.final |> select(-name) |> unique()

annotated |> group_by(gameId, trialNum, condition) |> select(flower) |> unique() |> tally() |> full_join(possible) |> 
  mutate(n=ifelse(is.na(n),0,n)) |> 
  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.2)+geom_smooth()+
          labs(y="Distinct flowers referenced per trial", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")
```

@@

annotated |> group_by(gameId, trialNum, condition, player, name) |> select(flower) |> unique() |> tally() |> full_join(d.rounds.final) |> 
  mutate(n=ifelse(is.na(n),0,n)) |>

  ggplot(aes(x=trialNum, y=n, color=condition))+geom_jitter(height=.25, alpha=.1)+geom_smooth()+
          labs(y="Distinct flowers referenced per player per trial", x="Trial Number")+
        theme(legend.position="none", legend.title=element_blank())+
        scale_x_continuous(limits=c(1,24), breaks=c(6,12,18,24))+
        scale_color_brewer(palette="Dark2")
```

@@

```{r, include=F,}

@@

 sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv"))|> 
    left_join(d.players) |> left_join(d.games.final) |> mutate(seen="game") |> 
   select(playerId, gameId, condition, text=normalized_span,referent, color, trialNum, seen) |> bind_cols(readRDS(here(annotation_path,'post_sbert.RData'))  %>% as_tibble())

sbert_post_test <- read_csv(here(annotation_path,"post_test_pre_sbert.csv")) |> 
  mutate(trialNum=25) |> 
 bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData'))  %>% as_tibble()) |> inner_join(d.games) |> 
   select(playerId, gameId, condition, text,referent=num, color, trialNum, seen, starts_with("V"))

sbert_all <- sbert_single |> union(sbert_post_test)
F_mat <- sbert_all %>% select(starts_with("...")) %>% as.matrix() #Features
M_mat <- sbert_all %>% 
  select(-starts_with("V")) %>% 
  mutate(feature_ind=row_number(),
         sixth=case_when(
           trialNum < 4 ~ 1,
          trialNum < 8 ~ 2,
          trialNum < 12 ~ 3,
          trialNum < 16 ~ 4,
          trialNum < 20 ~ 5,
          trialNum < 24 ~ 6,
          trialNum > 24 ~ 7
  ))
```
